---
title: "Statistical Inference Project Part-2"
author: "Shubhendu Dash"
date: "14/07/2020"
output:
  pdf_document: default
  html_document: default
---

## Introduction

This report aims to analyze the ToothGrowth data in the R datasets package. Per the course project instructions, the following items should occur:

1. Load the ToothGrowth data and perform some basic exploratory data analyses
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
4. State your conclusions and the assumptions needed for your conclusions.

## Analysis

**Load Libraries and ToothGrowth Data and investigate its structure.**

```{r}
library(datasets)
library(dplyr)
library(ggplot2)

# Load ToothGrowth Data

data("ToothGrowth")
str(ToothGrowth)
summary(ToothGrowth)
```

len : Tooth Length is numeric  
supp: Supplement is Factor consisting of OJ and VC  
dose: Dosage is numeric.  As it is just 0.5, 1 and 2, we will have to convert it to factor so that it is easier for analysis  

**Doing a quick plot to see if there are any patterns.**

```{r}

qplot(supp,len, data = ToothGrowth, colour = dose)
```

**There seems to be some patterns between supp and dose.  Let's use boxplot to investigate further.**

```{r}

ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# Boxplot for better view

g <- ggplot(ToothGrowth, aes(x=dose, y=len, colour=supp)) + geom_boxplot()
g
```

We have noticed:
1. There is a trend between dose and len.  
2. There is a trend between dose and supp.

### Hypothesis Testing No. 1

**1. Hypothesis Testing No.1: Overall impact of dosage on tooth growth.**
**2. Null Hypothesis - There is no effect of dosage on tooth growth.**
**3. Alternate Hypothesis - Higer doses produce greater tooth growth.** 

```{r}

# Arranging the dataset

dose05 <- filter(ToothGrowth, dose == 0.5)
dose1 <- filter(ToothGrowth, dose == 1)
dose2 <- filter(ToothGrowth, dose == 2)
```

**Comparing dosage of 1mg vs 0.5mg**

```{r}

# compare between dose of 0.5 and 1

t.test(dose1$len, dose05$len, alternative = "greater")
```

As P value is less than 0.001, it is highly significant and we can reject the null hypothesis

**Comparing dosage of 2mg vs 1mg**

```{r}

t.test(dose2$len, dose1$len, alternative = "greater")
```

### Conclusion No. 1

As P value is less than 0.001, it is highly significant and we can reject the null hypothesis. So in both casess, we reject the null hypothesis.  

**We accept the althernative: higher dosages do have a positive influence on toothgrowth.**

### Hypothesis Testing No. 2

**1. Hypothesis Testing No 2: Differences between supplements by dosage.**
**2. Null Hypothesis - There is no effect between supplement OJ and VC.**
**3. Alternate Hypothesis - OJ is better on toothgrowth than VC.**

We will do three tests and compare them based on dosages.

**Dosage: 0.5mg**

```{r}

OJdose05 <- filter(dose05, supp == "OJ")
VCdose05 <- filter(dose05, supp == "VC")
t.test(OJdose05$len, VCdose05$len, alternative = "greater")
```

P value is lower than 0.05.
We reject the null hypothesis.
We accept the alternative: OJ is a better supplement than VC at dosage of 0.5mg

**Dosage: 1mg**

```{r}

OJdose1 <- filter(dose1, supp == "OJ")
VCdose1 <- filter(dose1, supp == "VC")
t.test(OJdose1$len, VCdose1$len, alternative = "greater")
```

P value is lower than 0.05.
We reject the null hypothesis.
We accept the alternative: OJ is a better supplement than VC at dosage of 1mg

**Dosage: 2mg**

```{r}

OJdose2 <- filter(dose2, supp == "OJ")
VCdose2 <- filter(dose2, supp == "VC")
t.test(OJdose2$len, VCdose2$len, alternative = "greater")
```

P value is higher than 0.05.
We cannot reject the null hypothesis: at dosage of 2mg, there is no significant difference between the two supplement  

### Conclusion No. 2

**For dosages of 0.5mg and 1mg, there are significant differences: OJ is better than VC in promoting tooth growth.  At 2mg however, there is no significant differences.**

